---
title: "Averaging of biological replicates"
author: "Vladimir Kiselev"
date: "18 August 2015"
output:
  pdf_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(echo=FALSE)
```

```{r, include=FALSE}
source("functions.R")
```

# Averaging all bio replicates of PIP3

## Separated siINPP4(A+B) and siSHIP2 conditions

Looking at the file bio-reps-summary.pdf I decided to normalize all PIP3 data to the peak value in PTEN-sictrl, which is about 0.01 (in Golden Ration units). To do this I merge all data from the following files into one file and treat these different data as biological replicates:

2014_09_05B_c4_FULLn1

2014_09_19B_c4_FULLn2

2015_02_23B_c4_FULLn3

2014_12_18A_c4_full

2015_02_06B_c4_pten_ship2_ptenship2_n2

2015_04_04A_c4_wt_ship2_pten_ptenship2_N3

2014_12_12A_c4_wt_ab_pten_tc

2015_02_14B_c4_pten_ab_pab_tcN2_pabVSpabsN2

2015_03_06A_c4_pi34p2Trans_n3

2015_08_06A_c4_PTEN_crispr_n1

```{r}
rep1 <- get_bio_rep("2014_09_05B_c4_FULLn1", "P3_4")
rep1$rep <- "2014_09_05B_c4_FULLn1"
rep2 <- get_bio_rep("2014_09_19B_c4_FULLn2", "P3_4")
rep2$rep <- "2014_09_19B_c4_FULLn2"
rep3 <- get_bio_rep("2015_02_23B_c4_FULLn3", "P3_4")
rep3$rep <- "2015_02_23B_c4_FULLn3"
rep4 <- get_bio_rep("2014_12_18A_c4_full", "P3_4")
rep4$rep <- "2014_12_18A_c4_full"
rep5 <- get_bio_rep("2015_02_06B_c4_pten_ship2_ptenship2_n2", "P3_4")
rep5$rep <- "2015_02_06B_c4_pten_ship2_ptenship2_n2"
rep6 <- get_bio_rep("2015_04_04A_c4_wt_ship2_pten_ptenship2_N3", "P3_4")
rep6$rep <- "2015_04_04A_c4_wt_ship2_pten_ptenship2_N3"
rep7 <- get_bio_rep("2014_12_12A_c4_wt_ab_pten_tc", "P3_4")
rep7$rep <- "2014_12_12A_c4_wt_ab_pten_tc"
rep8 <- get_bio_rep("2015_02_14B_c4_pten_ab_pab_tcN2_pabVSpabsN2", "P3_4")
rep8$rep <- "2015_02_14B_c4_pten_ab_pab_tcN2_pabVSpabsN2"
rep9 <- get_bio_rep("2015_03_06A_c4_pi34p2Trans_n3", "P3_4")
rep9$rep <- "2015_03_06A_c4_pi34p2Trans_n3"
rep10 <- get_bio_rep("2015_08_06A_c4_PTEN_crispr_n1", "P3_4")
rep10$rep <- "2015_08_06A_c4_PTEN_crispr_n1"

d <- rbind(rep1, rep2)
d <- rbind(d, rep3)
d <- rbind(d, rep4)
d <- rbind(d, rep5)
d <- rbind(d, rep6)
d <- rbind(d, rep7)
d <- rbind(d, rep8)
d <- rbind(d, rep9)
d <- rbind(d, rep10)
```

Then I find a coefficient of normalisation for each replicate (so that it should be always 0.01 in PTEN-sictrl condition)

```{r, results='asis'}
coef <- d[Time == 1 & Condition == "sictrl" & Genotype == "PTEN", list(norm.coef = 0.01/Mean.GR), by = "rep"]
knitr::kable(coef)
```

Then I normalise each biological replicate by its coefficient.

```{r, include=FALSE}
setkey(coef, "rep")
setkey(d, "rep")
d <- d[coef]
d[ , Mean.GR.norm:=Mean.GR*norm.coef]
```

The resulting plot for PIP3 without PI3K inhibitor:

```{r, fig.width=8, fig.height=4}
p <- ggplot(d[Antagonist == "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() +
  theme(legend.text=element_text(size=5))
p
```

The resulting plot for PIP3 with PI3K inhibitor:

```{r, fig.width=8, fig.height=4}
p <- ggplot(d[Antagonist != "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() +
  theme(legend.text=element_text(size=5))
p
```

Now I remove the outlier experiment 2015_02_06B_c4_pten_ship2_ptenship2_n2 (it is an outlier for both cases with and without PI3K inhibitor). The resulting plots are the following:

```{r, fig.width=8, fig.height=4}
# remove the outlier
d <- d[rep != "2015_02_06B_c4_pten_ship2_ptenship2_n2"]

p <- ggplot(d[Antagonist == "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() +
  theme(legend.text=element_text(size=5))
p
```

```{r, fig.width=8, fig.height=4}
p <- ggplot(d[Antagonist != "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() +
  theme(legend.text=element_text(size=5))
p
```

Now I will average all of the replicates shown in the previous figure and plot the averaged data with error bars corresponding to biological noise:

```{r, fig.width=8, fig.height=4}
# average bio reps
d1 <- d[ , list(Bio.Mean.GR.norm = mean(Mean.GR.norm), Bio.SD.GR.norm = sd(Mean.GR.norm)),
  by = c("Component.Name.x", "Genotype", "Condition", "Antagonist", "Time")]

p <- ggplot(d1, aes(x = Time, y = Bio.Mean.GR.norm, color = Antagonist, group = Antagonist)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  geom_errorbar(aes(ymax = Bio.Mean.GR.norm + Bio.SD.GR.norm, ymin = Bio.Mean.GR.norm - Bio.SD.GR.norm), size = 0.5, width = 0.2) +
  theme_bw()
p
```

## siINPP4(A+B)&siSHIP2 condition

Note that there are still two experiments left in which we looked at double silencing of INPP4(A+B) and SHIP2:

2014_10_10A_c4_pabVSpabs

2015_02_14B_c4_pabVSpabsN2

```{r}
# add last 2 experiments and normalise them by siINPP4(A+B) (there is no sictrl in them)
rep11 <- get_bio_rep("2014_10_10A_c4_pabVSpabs", "P3_4")
rep11$rep <- "2014_10_10A_c4_pabVSpabs"
rep12 <- get_bio_rep("2015_02_14B_c4_pabVSpabsN2", "P3_4")
rep12$rep <- "2015_02_14B_c4_pabVSpabsN2"

t <- rbind(rep11, rep12)
```

I have not used them in the previous section because they do not have PTEN-sictrl condition and would not be able to normalise them. But now, looking the last figure, I can normalise them by PTEN-siINPP4(A+B) condition (all other bio replicates are quite consistent in this point). The PIP3 concentration at this point should be:

```{r, echo=TRUE}
d1[Genotype == "PTEN" & Condition == "siINPP4(A+B)" & Time == 1, Bio.Mean.GR.norm]
```

So, first I normalise these two experiments by this time point:

```{r, results='asis'}
coef <- t[Time == 1 & Condition == "siINPP4(A+B)" & Genotype == "PTEN", list(norm.coef = 0.00745414134784178/Mean.GR), by = "rep"]
knitr::kable(coef)
```

```{r, include=FALSE}
setkey(coef, "rep")
setkey(t, "rep")
t <- t[coef]
t[ , Mean.GR.norm:=Mean.GR*norm.coef]
```

And then merge these two replicates with the ones defined before.

```{r}
# merge all bio reps
all.rep <- rbind(d, t)
```

Now the total data set looks like:

```{r, fig.width=8, fig.height=4}
p <- ggplot(all.rep[Antagonist == "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() + 
  theme(legend.text=element_text(size=5))
p
```

```{r, fig.width=8, fig.height=4}
p <- ggplot(all.rep[Antagonist != "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() +
  theme(legend.text=element_text(size=5))
p
```

## Final plot of PIP3 data

And averaging of all biological replicates provides a final plot containing all PIP3 data (I will be working with it in the modeling part):

```{r, fig.width=8, fig.height=4}
# average bio reps
d2 <- all.rep[ , list(Bio.Mean.GR.norm = mean(Mean.GR.norm), Bio.SD.GR.norm = sd(Mean.GR.norm)),
        by = c("Component.Name.x", "Genotype", "Condition", "Antagonist", "Time")]

write.csv(d2, file = "all-data/bio-rep-av/p3_4.csv", row.names = F, quote = F)

p <- ggplot(d2, aes(x = Time, y = Bio.Mean.GR.norm, color = Antagonist, group = Antagonist)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  geom_errorbar(aes(ymax = Bio.Mean.GR.norm + Bio.SD.GR.norm, ymin = Bio.Mean.GR.norm - Bio.SD.GR.norm), size = 0.5, width = 0.2) +
  theme_bw()
p
```

# Averaging all bio replicates of PI(3,4)P2

The same procedure as in the previous chapter can be done for PI(3,4)P2 lipids.

## Separated siINPP4(A+B) and siSHIP2 conditions

Looking at the file bio-reps-summary.pdf I decided to normalize all PI(3,4)P2 data to the peak value in PTEN-sictrl. To do this I merge all data from the following files into one file and treat these different data as biological replicates:

2014_09_05A_c18_FULLn1

2014_09_19A_c18_FULLn2

2015_02_23A_c18_FULLn3

2014_12_12B_c18_wt_ab_pten_tc

2015_02_14A_c18_pten_ab_pab_tcN2_pabVSpabsN2

2015_03_06B_c18_pi34p2Trans_n3

2015_08_06B_c18_PTEN_crispr_n1

```{r}
rep1 <- get_bio_rep("2014_09_05A_c18_FULLn1", "SA_P2_aldehyde_34")
rep1$rep <- "2014_09_05A_c18_FULLn1"
rep2 <- get_bio_rep("2014_09_19A_c18_FULLn2", "SA_P2_aldehyde_34")
rep2$rep <- "2014_09_19A_c18_FULLn2"
rep3 <- get_bio_rep("2015_02_23A_c18_FULLn3", "SA_P2_aldehyde_34")
rep3$rep <- "2015_02_23A_c18_FULLn3"
rep4 <- get_bio_rep("2014_12_12B_c18_wt_ab_pten_tc", "SA_P2_aldehyde_34")
rep4$rep <- "2014_12_12B_c18_wt_ab_pten_tc"
rep5 <- get_bio_rep("2015_02_14A_c18_pten_ab_pab_tcN2_pabVSpabsN2", "SA_P2_aldehyde_34")
rep5$rep <- "2015_02_14A_c18_pten_ab_pab_tcN2_pabVSpabsN2"
rep6 <- get_bio_rep("2015_03_06B_c18_pi34p2Trans_n3", "SA_P2_aldehyde_34")
rep6$rep <- "2015_03_06B_c18_pi34p2Trans_n3"
rep7 <- get_bio_rep("2015_08_06B_c18_PTEN_crispr_n1", "SA_P2_aldehyde_34")
rep7$rep <- "2015_08_06B_c18_PTEN_crispr_n1"

d <- rbind(rep1, rep2)
d <- rbind(d, rep3)
d <- rbind(d, rep4)
d <- rbind(d, rep5)
d <- rbind(d, rep6)
d <- rbind(d, rep7)
```

To find the averaged concentration in PTEN-sictrl condition I simply average the data in this point (there are no really bad outliers in PI(3,4)P2 data, so it is possible to do it):

```{r, echo=TRUE}
d[Time == 1 & Condition == "sictrl" & Genotype == "PTEN", list(mean(Mean.GR))]
```

Then I find a coefficient of normalisation for each replicate (so that it should be always 0.0765314027687824 in PTEN-sictrl condition)

```{r, results='asis'}
coef <- d[Time == 1 & Condition == "sictrl" & Genotype == "PTEN", list(norm.coef = 0.0765314027687824/Mean.GR), by = "rep"]
knitr::kable(coef)
```

Then I normalise each biological replicate by its coefficient.

```{r, include=FALSE}
setkey(coef, "rep")
setkey(d, "rep")
d <- d[coef]
d[ , Mean.GR.norm:=Mean.GR*norm.coef]
```

The resulting plot for PIP3 without PI3K inhibitor:

```{r, fig.width=8, fig.height=4}
p <- ggplot(d[Antagonist == "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() +
  theme(legend.text=element_text(size=5))
p
```

The resulting plot for PIP3 with PI3K inhibitor:

```{r, fig.width=8, fig.height=4}
p <- ggplot(d[Antagonist != "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() +
  theme(legend.text=element_text(size=5))
p
```

Now I will average all of the replicates shown in the previous figure and plot the averaged data with error bars corresponding to biological noise:

```{r, fig.width=8, fig.height=4}
# average bio reps
d1 <- d[ , list(Bio.Mean.GR.norm = mean(Mean.GR.norm), Bio.SD.GR.norm = sd(Mean.GR.norm)),
  by = c("Component.Name.x", "Genotype", "Condition", "Antagonist", "Time")]

p <- ggplot(d1, aes(x = Time, y = Bio.Mean.GR.norm, color = Antagonist, group = Antagonist)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  geom_errorbar(aes(ymax = Bio.Mean.GR.norm + Bio.SD.GR.norm, ymin = Bio.Mean.GR.norm - Bio.SD.GR.norm), size = 0.5, width = 0.2) +
  theme_bw()
p
```

## siINPP4(A+B)&siSHIP2 condition

Note that there are still two experiments left in which we looked at double silencing of INPP4(A+B) and SHIP2:

2014_10_10B_c18_pabVSpabs

2015_02_14A_c18_pabVSpabsN2

```{r}
# add last 2 experiments and normalise them by siINPP4(A+B) (there is no sictrl in them)
rep8 <- get_bio_rep("2014_10_10B_c18_pabVSpabs", "SA_P2_aldehyde_34")
rep8$rep <- "2014_10_10A_c4_pabVSpabs"
rep9 <- get_bio_rep("2015_02_14A_c18_pabVSpabsN2", "SA_P2_aldehyde_34")
rep9$rep <- "2015_02_14B_c4_pabVSpabsN2"

t <- rbind(rep8, rep9)
```

I have not used them in the previous section because they do not have PTEN-sictrl condition and would not be able to normalise them. But now, looking the last figure, I can normalise them by PTEN-siINPP4(A+B) condition (all other bio replicates are quite consistent in this point). The PIP3 concentration at this point should be:

```{r, echo=TRUE}
d1[Genotype == "PTEN" & Condition == "siINPP4(A+B)" & Time == 1, Bio.Mean.GR.norm]
```

So, first I normalise these two experiments by this time point:

```{r, results='asis'}
coef <- t[Time == 1 & Condition == "siINPP4(A+B)" & Genotype == "PTEN", list(norm.coef = 0.146433426237419/Mean.GR), by = "rep"]
knitr::kable(coef)
```

```{r, include=FALSE}
setkey(coef, "rep")
setkey(t, "rep")
t <- t[coef]
t[ , Mean.GR.norm:=Mean.GR*norm.coef]
```

And then merge these two replicates with the ones defined before.

```{r}
# merge all bio reps
all.rep <- rbind(d, t)
```

Now the total data set looks like:

```{r, fig.width=8, fig.height=4}
p <- ggplot(all.rep[Antagonist == "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() + 
  theme(legend.text=element_text(size=5))
p
```

```{r, fig.width=8, fig.height=4}
p <- ggplot(all.rep[Antagonist != "-"], aes(x = Time, y = Mean.GR.norm, color = rep, group = rep)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  theme_bw() +
  theme(legend.text=element_text(size=5))
p
```

## Final plot of PI(3,4)P2 data

And averaging of all biological replicates provides a final plot containing all PI(3,4)P2 data (I will be working with it in the modeling part):

```{r, fig.width=8, fig.height=4}
# average bio reps
d2 <- all.rep[ , list(Bio.Mean.GR.norm = mean(Mean.GR.norm), Bio.SD.GR.norm = sd(Mean.GR.norm)),
        by = c("Component.Name.x", "Genotype", "Condition", "Antagonist", "Time")]

write.csv(d2, file = "all-data/bio-rep-av/pi34p2.csv", row.names = F, quote = F)

p <- ggplot(d2, aes(x = Time, y = Bio.Mean.GR.norm, color = Antagonist, group = Antagonist)) +
  geom_point() +
  geom_line() +
  facet_grid(Genotype ~ Condition) +
  geom_errorbar(aes(ymax = Bio.Mean.GR.norm + Bio.SD.GR.norm, ymin = Bio.Mean.GR.norm - Bio.SD.GR.norm), size = 0.5, width = 0.2) +
  theme_bw()
p
```
